---
title: "data_wrangling"
author: "Chelsea Christie"
date: "October 16, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data wrangling notes

see:
http://stat545.com/block006_care-feeding-data.html
https://bookdown.org/rdpeng/exdata (separates out many functions so it's easy to find individual ones)

Start by loading my favorite packages

```{r}
library(tidyverse)
library(psych)
library(foreign)
library(forcats)
library(ggplot2)
```

Start by exploring what is in the dataset:

*only print the dataset to screen if the tidyverse is loaded (just get a tibble and not the entire thing!) OR can use as_tibble() to show something as a tibble

head(): just shows the beginning 
tail(): just shows the end

str(): displays the structure of an object and will provide a sensible description of almost anything




Some basic functions in dplyr (part of the tidyverse):

select(): return a subset of the columns of a data frame, using a flexible notation

filter(): extract a subset of rows from a data frame based on logical conditions

arrange(): reorder rows of a data frame

rename(): rename variables in a data frame

mutate: add new variables/columns or transform existing variables

summarise / summarize: generate summary statistics of different variables in the data frame, possibly within strata


