---
title: "data_wrangling"
author: "Chelsea Christie"
date: "October 16, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data wrangling notes

see:
http://stat545.com/block006_care-feeding-data.html
https://bookdown.org/rdpeng/exdata (separates out many functions so it's easy to find individual ones)

Start by loading my favorite packages

```{r}
library(tidyverse)
library(psych)
library(foreign)
library(forcats)
library(ggplot2)
```

Start by bringing in the dataset:
see:
http://stat545.com/block026_file-out-in.html
http://r4ds.had.co.nz/data-import.html

If it's an excel file, use 'read_csv()'


exploring what is in the dataset & basic functions:

http://r4ds.had.co.nz/transform.html
http://stat545.com/block009_dplyr-intro.html

*only print the dataset to screen if the tidyverse is loaded (just get a tibble and not the entire thing!) OR can use as_tibble() to show something as a tibble

head(): just shows the beginning 
tail(): just shows the end

str(): displays the structure of an object and will provide a sensible description of almost anything

ncol() or nrow(): gives you # of columns or rows

select(): pick variables by their names

filter(): extract a subset of rows from a data frame 

arrange(): reorder rows of a data frame

rename(): rename variables in a data frame

mutate(): add new variables/columns or transform existing variables

summarise / summarize(): generate summary statistics of different variables in the data frame (collapse many value down to 1 single summary statistic)

group_by(): changes the scope of each function from operating on the entire dataset to operating on it group-by-group. 

** a nice tidy way to write longer chunks of code is with the pipe ( %>% ctr + shift + m). Think of the pipe as 'then' so you can write code with the order you want the computer to do it.


