---
title: "ecoscope_workshop_notes"
author: "Chelsea Christie"
date: "January 23, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Jan. 23, 2017

```{r}


rm(months_chelsea_has_lived)
rm(x,y)
seq(1,100)
months_chelsea_has_lived <- 314

data("mtcars")
mtcars

library(tidyverse)
View(mtcars)
View(x= mtcars, title = "Motor Trend")

help(View)
?View

str(mtcars)
rownames(mtcars)
dim(mtcars)

summary(mtcars)

library(dplyr)
library(help="dplyr")

sapply(mtcars, mean)
min(mtcars$mpg)
quantile(mtcars$mpg, probs = .5)
quantile(mtcars$mpg, probs = c(0, 1, .7, .9))
table(mtcars$gear)
xtabs(~cyl, data = mtcars)
xtabs(~cyl+gear, data = mtcars)
lm(mpg~ cyl + disp, data = mtcars)
# above line gives the linear regression showing the relationship between number of cylinders and miles per gallon. The criterion variables get listed first.

sapply(mtcars, quantile, probs = c(.25, .75))

sapply(mtcars[,c("mpg", "disp")], quantile, probs = c(.25, .75))

```

Objects can not start with underscores or numbers and they are case-senstive. When creating an object that is composed of letters you have to put "" around the word(s). []s tell you which place the object is within the vector. So if you have a list of 10 numbers in chronological order, [4] is 4. To clear console = "ctrl + l ". 

# Getting help with functions
To get more information about a function, use "help(function name)" or "?function name". Have to respect the order of arguments within functions unless you specifically name the argument (e.g. title comes second in view but you could put it first if you type "title = whatever the title is").

-to get number of rows and columns = dim(data file name) OR ncol() and nrow()

# some notes about vectors

-vectorization means that when you do something to one thing in the dataframe it will happen to every item in the vector. So you can easily square or transform every point in the dataframe.

mean(), sd() = standard deviation, 
sapply(dataframe, whatever you want applied) -e.g. sapply(mtcars, mean) will give you the mean for every variable within the vector
lm() = linear model, use for regression

# To subset to specific columns, can use square brackets e.g. [,c("variable one", "variable 2")]

## Jan. 25, 2017 - ggplot2

```{r}

library(tidyverse)
library(ggplot2)

?mpg

p <- ggplot(data = mpg,
       aes(x = displ, y = cty, color = class)) +
  geom_point()

p

#ggsave("plot1.pdf")

# can also pick the colors you want

p2 <- ggplot(data = mpg,
       aes(x = displ, y = cty)) +
  geom_point(color = 'skyblue')

p2

# to see all colours - just google it or can do 

#colors()

# non data elements get changed through the 'theme' function

?theme


ggplot(mpg, aes(x = cty)) +
  geom_histogram(bins = 20)


ggplot(mpg, aes(x = factor(cyl), y = cty)) +
  geom_boxplot()

# to order the elements in the graph - use 'reorder'  - ex. below

ggplot(mpg, aes(x = reorder(drv, cty, median),
                y = cty)) +
  geom_boxplot()

# or can tell it exactly the order to put things in by using 'levels =c()

ggplot(mpg, aes(x = factor(drv,levels = c("f", "r", "4")),
                y = cty)) +
  geom_boxplot()

ggplot(economics,
       aes(y = unemploy, x = date)) +
  geom_line() +
  geom_smooth()

# to get rid of the confidence interval on the geom_smooth() - use se = FALSE
# or to add a regression line - use method ="lm"

ggplot(economics,
       aes(y = unemploy, x = date)) +
  geom_line(color = "red") +
  geom_smooth(se=FALSE, method ="lm", color = "black")

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = factor(cyl))) +
  geom_smooth(aes(y = cty), se = FALSE) +
  labs(title = "Highway fuel consumption",
       x= "displacement",
       y= "highway mpg")

# to add titles use 'ggtitle()' as another layer
# to change labels use 'labs() as another layer

library(gapminder)
?gapminder

ggplot(gapminder, aes(x = year, y = lifeExp, group = country)) +
  geom_line(aes(color = factor(continent), alpha = 0.5))

# alpha - makes things more or less transparent

ggplot(mpg, aes(displ, hwy, color = factor(cyl))) +
  geom_point() +
    facet_wrap(~manufacturer)

#use position to play around with where your datapoints go. e.g position dodge so bar charts don't overlap, or position nudge to make sure label points are not over top of the data points

```

# Jan 27 - dyplr functions - organizing data

```{r}

library(tidyverse)

###### Select - for choosing columns 

# to look at just two variables - use select()

select(mtcars, wt, mpg)

# or just variables that start with 'd'

select(mtcars, starts_with("d"))

# can also use 'ends_with or contains
# can also put one variable at the front by listing it first and then everything

select(mtcars,wt, everything())


##### Filter - for choosing rows or observations

filter(mtcars, cyl ==8)

# one equals sign means you're assigning an argument to a function (like using <- ) but use a double equals sign to demonstrate that you want values to be equal to something. So in the above example I am saying return all cylinders that are 8 and NOT making a new object - like one = sign would suggest.

# can use other logical arguments- like less than 

filter(mtcars, cyl <8)

# if you want to keep the row names in the output you have to use add_rownames

filter(rownames_to_column(mtcars), cyl < 8)

## to combine filter & select
# ex. choose all cars with 8 cylinders and

df <- filter(mtcars, cyl == 8)

select(df, disp, wt)

?mtcars

## Exercise 

filter(mtcars, am == 1)

cars_with_5_gears_or_manual_transmission <- filter(mtcars, gear == 5 | am == 1)

select(cars_with_5_gears_or_manual_transmission, mpg, wt)


select(filter(mtcars, am == 1), qsec)


###### The pipe!!!

mtcars %>% 
  filter(am == 1) %>% 
  select(qsec)


## think of the pipe as 'insert' - where everything to the left of the pipe gets inserted into the first argument of the next function

#### Arrange - sorting your dataframe 

arrange(mtcars, mpg)
arrange(mtcars, desc(mpg))

mtcars %>% 
  arrange(cyl, carb)

mtcars %>% 
  arrange(desc(mpg & disp))

?nycflights13

install.packages("nycflights13")
library(nycflights13)

?nycflights13

str(flights)
?flights

glimpse(flights)

# when saying you DON'T want certain variables you can use the minus sign in front of it.

flights %>% 
  filter(day == 2, month == 1) %>% 
  select(-month, -year, -day) %>% 
  glimpse()

    
# which flights are the most delayed?

flights %>% 
  arrange(desc(arr_delay)) %>% 
  head()

# which flights caught up the most time over flights (looking at difference between departure delay and arrival delay)

flights %>% 
  arrange(desc(dep_delay - arr_delay)) %>% 
  select(carrier)

# transmute will mutate old variables into new ones AND delete the old ones

mtcars %>% 
  transmute(disp_1 = disp/61.0237)

mtcars %>% 
  mutate(wt_in_tonnes = wt/2204.6)

#### summmarize function - compute things for a variable

mtcars %>% 
  summarize(mean(mpg), med_wt = median(wt))

## group by - used to do operations for certain parts of your data

mtcars %>% 
  group_by(cyl) %>% 
  summarize(mean_mpg = mean(mpg), med_wt= median(wt))

# can also group by more than one variable

mtcars %>% 
  group_by(am, cyl) %>% 
  summarize(mean_mpg = mean(mpg), med_wt= median(wt))

mtcars %>% 
  group_by(am, vs) %>% 
  summarize(min(disp), max(disp))


?flights

flights %>%
  mutate(speed_mph = distance/(air_time/60)) %>% 
  select(speed_mph, flight) %>% 
  arrange(desc(speed_mph)) %>% 
  head()

### another tip about pipes

boxplot(mpg~cyl, data = filter(mtcars, cyl < 8))

mtcars %>% 
  filter(cyl <8) %>% 
  boxplot(mpg~cyl, data =.)

?boxplot

## if you want to use a pipe but the function you are using does not use data as the first argument - you can use 'data =.' to indicate that the data you want to use is the one you listed higher up in the pipe.

# exercise: which destinations have the hightest average delays
?flights

flights %>% 
  group_by(dest) %>% 
  summarize(delay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(delay))

# na.rm = NA remove. so missing values will be excluded when calculating the average.

```

