---
title: "general_rRstudio_notes"
author: "Chelsea Christie"
date: "October 1, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chelsea's notes from Stat545 about how to use R Studio

### Basic R notes

see:
http://stat545.com/block002_hello-r-workspace-wd-project.html

For more tips on R markdown:
http://stat545.com/block007_first-use-rmarkdown.html
or
http://r4ds.had.co.nz/r-markdown.html

Start by loading my favorite packages

```{r}
library(tidyverse)
library(psych)
library(foreign)
library(forcats)
library(ggplot2)
```


```{r}
# <- ## makes an assignment (kind of like giving something a name/short hand)
objects() ## tells me what objects are in my workspace 
# ls() ## also tells me what objects are in my workspace
# rm() ## will remove an object
# rm(list = ls()) ## will remove all objects in workspace (can also do with broom in the environment)

# (name <- whatever I'm naming/assigning ) putting an assignment in brackets will cause it print at the same time

# q() = quit Rstudio 

# getwd() = shows which working directory I"m "working in" and importantly where all the stuff I'm doing will get saved to

# Rstudio projects keep all the files associated with a project organized together – input data, R scripts, analytical results, figures

# It's always a good idea to test run things using fake data. To make a fake dataset:

(A <- data.frame(
        c1 = c('A', 'A', 'A', 'A', 'A', 'B', 'B'),
        c2 = c('a', 'a', 'a', 'b', 'b', 'c', 'd'),
        c3 = c(1, 3, 1, 1, 2, 2, 1)))

# to make tables pretty, use:

# knitr::kable()

# For example: 

(knitr::kable(A))

```

Some more general Rstudio information:

%>%: the “pipe” operator is used to connect multiple verb actions together into a pipeline. Once you travel down the pipeline with %>%, the first argument is taken to be the output of the previous element in the pipeline.