---
title: "R Visualization with Duncan Murdoch"
author: "Chelsea Christie"
date: "March 3, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Start by loading favorite packages

```{r}
library(tidyverse)
library(psych)
library(foreign)
library(forcats)
library(ggplot2)
library(rgl)
library(dagitty)
library(ggm)
library(BayesFactor)
```

## March 3

Slides available on his website: 

http://www.stats.uwo.ca/faculty/murdoch/Calgary2018/

-provides history of different r graphics packages

-ggplot2 came out in 2007 - Hadley Wickham used Wilkinson's ideas about grammar graphics
----ggplot stands for grammar of graphics plot

-rgl is a package for drawing 3D interactive plots (developed by Murdoch & Daniel Adler)
----rgl stands for r graphics language

-htmlwidgets allows you to connect to javascript libraries of graphics including cool mapping graphics (e.g. package called leaflet)

-rcolorbrewer is a good package for finding colour palettes that are aesthetically pleasing and friendly to colour blind people 

## Notes about ggplot2

** Great to buy Hadley Wickham's book: ggplot2: Elegant Graphics for Data Analysis 
Or R graphics Cookbook

data: your data

aesthetics: things like colour, shape, size, etc. (How your data maps to the display)

layers or geoms: shapes that represent the data, produced by various functions with names like geom_X()

Other ways to modify the graph:

scales: how the data maps to the plot
transformations: changes to the data
coordinate systems
facets: multiple plots together
visual themes: overall styles of aesthetics 

** to see all possible geoms (different plot options):

ls("package:ggplot2", pattern = "^geom_")

## Practicing plots using mtcars dataset

```{r}
# learn about the dataset
?mtcars

# create a basic plot (first argument = data, then first argument for aes is x axis and second argument for aes is y axis)
ggplot(mtcars, aes(x =wt, y =mpg))

# above plot has no layers (geoms) so it's blank but the coordinate system is outlined

# now make it a scatterplot
ggplot(mtcars, aes(wt, mpg)) + geom_point()

# can save the base part of the plot (the first part of the plot) and then assign it a name so you can easily add to it later

base_plot_wt_mpg <- ggplot(mtcars, aes(wt, mpg)) + geom_point()

base_plot_wt_mpg + aes(colour = qsec)

base_plot_wt_mpg + geom_point(colour = "green")

# can show two geoms at once

base_plot_wt_mpg + geom_density2d(colour = "white")

# to pretty up the previous plot, I might want to remove the background grid lines. Can use the theme() function in order to change the general look of things. 

base_plot_wt_mpg + geom_density2d(colour = "white") + geom_point() +
theme(panel.grid = element_blank()) +
xlab("Weight") + ylab("Miles per gallon")

# some other theme tips -- can use theme_dark() to get a dark background 
# last_plot() can also be used to pull up everything from the previous plot & then try out an additional thing

# faceting is a way to show multiple plots, can use facet_wrap to show multiple plots beside each other, or facet_grid to show multiple plots with the rows and columns to be the same variable

# ~ is used to remember a formula 

```

## some other ggplot2 notes

Other useful functions:

labs(), ggtitle(): modify labels on plots.
annotation_*(): add notes to plots
coord_*() functions change the display from the default Cartesian coordinates, e.g. to map projections
draw_key*() functions draw elements in legends

## Now moving onto 3D plots using rgl

```{r}

# can use rglwidget() to view the plot

library(rgl)
options(rgl.useNULL = TRUE)
plot3d(iris, col = as.numeric(iris$Species), type = "s")
rglwidget()

cybera.sy.gy.za.ca

```

